---
title: "Code"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'code.html'))})
output:
  html_document:
    theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction of R

I am using R Studio for class examples which is more user friendly. The computation core is still R. It is okay if you want to use and familiar with others, but I may not provide full assistance. R can be downloaded from *https://www.r-project.org/*. RStudio is from *https://rstudio.com/*.

```{r}
# setwd("G:/Desktop/rsite/stat464") # set working directory
# install.packages("MPV")
library(MPV) # data in the appendix
# install.packages("scatterplot3d") # Install
library("scatterplot3d") # load
# install.packages("installr")
library(installr) 
#install.packages("xlsx")
library(readxl)
p2.7
head(p2.7)
```

# Simple Linear Regression
## Regression and Inference
```{r}
example21 = read_excel("data/Chapter 2/Examples/data-ex-2-1 (Rocket Prop).xls") # read .xls file from local directory
colnames(example21) = c("Observation","Shear_Strength_psi","Age_of_Propellant_weeks")
example21
head(example21)
attach(example21)
y = example21$Shear_Strength_psi
x = example21$Age_of_Propellant_weeks
n = length(x)
plot(Age_of_Propellant_weeks,Shear_Strength_psi,pch = 16)
```

Our model is
\[
Y = \beta_0+\beta_1 X+\epsilon.
\]

Here, $Y$ is *Shear_Strength_psi* and $X$ is *Age_of_Propellant_weeks*. We can compute what we want directly from definitaions
\[
S_{xx} = \sum_{i=1}^n(x_i-\bar{x})^2\\
S_{xy} = \sum_{i = 1}^ny_i(x_i-\bar{x})=\sum_{i=1}^nx_iy_i-\bar{x}\sum_{i=1}^ny_i\\
\hat{\beta}_1 = \frac{S_{xy}}{S_{xx}}\\
\hat{\beta}_0 = \bar{y}-\hat{\beta}_1\bar{x}.
\]

``` {r}
sxx = sum(x^2)-sum(x)^2/n;sxx
sxy = sum(x*y)-sum(x)*sum(y)/n;sxy
beta1 = sxy/sxx;beta1
beta0 = mean(y)-beta1*mean(x)
c(beta0,beta1)
error = y-(beta0+beta1*x)
plot(error);hist(error)
```

We can use R packages to do most of the work in the future. Most softwares have packages for linear model. In *lm* function, we have everything we want in the output summary. 

``` {r}
model1 = lm(Shear_Strength_psi ~ Age_of_Propellant_weeks,data = example21)
summary(model1)
```

For example, we know the variance of slope $\beta_0$ is
\[Var(\hat{\beta_1})=\frac{\sigma^2}{S_{xx}}\].
The test statistic of $\beta_0$ is 
\[
t_0 = \frac{\hat{\beta_0}-\beta_{00}}{\sqrt{MS_{Res}(1/n+\bar{x}^2/S_{xx})}}.
\]
*p-value* is given by
\[P(T>t_0).\]
Let's do it by hand.

``` {r}
sst = sum((y-mean(y))^2);sst
ssres = sum(error^2);ssres
varBeta1 = (ssres/(n-2))/sxx;sqrt(varBeta1) # variance of \beta_1
sdOfBeta1 = sqrt(ssres/(n-2)*(1/sxx))
sdOfBeta0 = sqrt(ssres/(n-2)*(1/n+mean(x)^2/sxx))
t0 = (beta0-0)/sdOfBeta0;t0
pt(-t0,n-2)
```

Similarly, *lm* returns everything else as we want. (We will meet *Adjusted R-squared* later).

Let $\alpha = 0.05$, the $(1-\alpha)100\%=95\%$ confidence interval (C.I.) of $\beta_1$ is $(-43.22 -31.08)$, given by
``` {r}
c(beta1+qt(0.025,n-2)*sdOfBeta1,beta1-qt(0.025,n-2)*sdOfBeta1)
```

## Prediction
We predict the average response at point $x_0$ with its corresponding confidence interval. For all $x_i,i=1,...,n$, we can create a confidence band.

``` {r}
yhat = beta0+beta1*x
plot(x,y,pch=16)
lines(x,yhat)
x0 = 15;yhat0 = beta0+beta1*x0
ciPoint = c(yhat0-qt(0.025,n-2)*sqrt(ssres/(n-2)*(1+1/n+(x0-mean(x))^2/sxx)),
            yhat0+qt(0.025,n-2)*sqrt(ssres/(n-2)*(1+1/n+(x0-mean(x))^2/sxx)))
points(x0,yhat0,pch=2,cex=2);points(c(x0,x0),ciPoint,pch = 8)
yhatSort = beta0+beta1*sort(x)
ciBand = cbind(yhatSort-qt(0.025,n-2)*sqrt(ssres/(n-2)*(1/n+(sort(x)-mean(x))^2/sxx)),
            yhatSort+qt(0.025,n-2)*sqrt(ssres/(n-2)*(1/n+(sort(x)-mean(x))^2/sxx)))
ciPredBand = cbind(yhatSort-qt(0.025,n-2)*sqrt(ssres/(n-2)*(1+1/n+(sort(x)-mean(x))^2/sxx)),
            yhatSort+qt(0.025,n-2)*sqrt(ssres/(n-2)*(1+1/n+(sort(x)-mean(x))^2/sxx)))
plot(x,y,pch=16)
lines(x,yhat,lwd = 2)
lines(sort(x),ciBand[,1],type = 'l',lty = 'dashed',lwd = 2)
lines(sort(x),ciBand[,2],type = 'l',lty = 'dashed',lwd = 2)
lines(sort(x),ciPredBand[,1],type = 'l',lty = 'dotted',lwd = 2)
lines(sort(x),ciPredBand[,2],type = 'l',lty = 'dotted',lwd = 2)
```

Of course, we have simple way.

``` {r}
confint(model1)
newData = data.frame(Age_of_Propellant_weeks = sort(Age_of_Propellant_weeks))
bandOfConf = predict(model1, newdata = newData, interval = 'confidence')# "confidence"
bandOfPred = predict(model1, newdata = newData, interval = 'prediction')# "confidence"
xOrdered = newData$Age_of_Propellant_weeks
plot(x,y)
abline(model1, col = "red")
lines(xOrdered,bandOfConf[,2],type = 'l',lty = 'dashed',lwd = 2)
lines(xOrdered,bandOfConf[,3],type = 'l',lty = 'dashed',lwd = 2)
lines(xOrdered,bandOfPred[,2],type = 'l',lty = 'dotted',lwd = 2)
lines(xOrdered,bandOfPred[,3],type = 'l',lty = 'dotted',lwd = 2)
```

If we want to estimate the line through the origin, we can do the following

