---
title: "Notes"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
output:
  html_document:
    toc: true
    toc_float: TRUE
    theme: spacelab
---

```{r,echo=F, message=F}
library(dplyr)
library(kableExtra)
```

# About

These are class notes and R code for Professor Ge Zhao's STAT-464 : Applied Regression Analysis for Fall term 2021 at Portland State University. 

# D1 : Monday September 27, 2021

- will ask around for a bigger room 

- assume we have basic understanding of R 

## Introduction : Chapter 1 

- We want to know if the data matches (1-1), (1-Many), (Many - Many) and such 

- We call the data we are interested in the response $Y$ or sometimes the independet variable 

  - want to see what happens to y and the change of Y (predict Y)
  
  - examples : price of bitcoin, or price of house 

- We call the $X$ the predictor(s)

- want to use prediction to predict a response 

  - view of the model 
  
  - annalyze a model 
  
  - example (bitcoin) : day 1 = \$50, day 2 = \$100, day 3 = \$1,000
  
- math model is a statistical model, and to firgue out $Y=f(x)+\epsilon$ where $\epsilon$ is random error

  - in math model everything is fixed except for the $\epsilon$
  
  - this is a very general problem (tells almost nothing)
  
```{r, echo=F}
X <- c("continuous", "categrorical : ordinal data and nominal data")
Y <- c("continuous", "categorical : ordinal data and nominal data")

data.frame(X,Y) %>% kable() %>% kable_paper
```
- ordinal data (can be ordered) and nominal data (different types of data)"

- in regression class most of the time we are focuses on X=Y : Continuous vs Continuous 

- in experimental design we are focused on Y=X : Continuous vs Categorical 

- special case (in later chapters) Y=X : Ordinal vs. continuous 

- later (in other classes) we may see (double line) experimental design looks at every combination of categorical vs. categrorical .

### Summary / Thoughts

We are focusing on data, and relations between $X$ and $Y$. 

Regressions would have continuous vs. continuous. This reminds me graphically of physics problems. How far does something go in a certain amount of time. 

Experimental design would have a continuous $Y$ vs. a Categorical $X$. Graphically I think of histograms. 

Some of the special cases I would need to think about. 

